---
title: "Covid91 vaccine study Final2025"
author: "Deztinee Saldana"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
M = filter(FinalData, sex =="M")
F = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Coding details 

In order for the data to be evaluated efficiently, it was important to seperate the variables into the four categories that they belonged to. I made four variables called "M", "F", "gay", and "drug users". They were isolated in order to see whether or not they should take the Covid-19 Vaccine. This was so I could create the effectiveness intervals. They were stored in seperate variables. 

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

### Barplot

```{r}
dd2 <- M %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot shows the population of male people and the relationship between catching Covid, whether or not they took the vaccine or had a placebo. 

The first graph shows the counts of the people who got covid or if they were okay and if they took the placebo or not. The second graph shows the percentage of people who had the placebo or the covid-19 vaccine and whether if they had covid or not. They are all split up in groups by if they got covid-19 or not and if they had the placebo or the Covid-19 vaccine. 

Of the people who caught Covid-19, the people who took the placebo were plot as a big bar being approximately around 75% and the people who had the vaccine were significantly small, being approximately around 25%. Of the people who were OK, it was split 50/50 between people who took the placebo vs. the people who had the vaccine. 

The codes that were used to create this barplot start with the code block that begins with dd2. dd2 appars to be seen where the environment pane and it contains 4 variables. These variables include treatment, infected, count, and prcent(which stands for percent). The second group is named the basicC. It is visible in the environment area and it contains a list of 11 variables. This is in order to create the first plot. The basicCC is the last group of codes and it is also in the files pane. It contains 11 variables that are helpful when creating the percentage plot, which is the second plot. 

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=M)
rowPerc(table1)
colPerc(table1)
```

These numerical results show the actual percentages of the males that got covid-19 after they took the placebo or the vaccine. 

The males that got covid-19 after taking the placebo was a huge value of 73.80%, compared to the percentage of people who got covid-19 after taking the vaccine (26.20%). Of the males that were okay, 50.01% of the men who took the placebo were fine, and 49.99% of the men who took the vaccine were fine. This shows that the covid-19 vaccine is effective in males if only 26.20% of the people who took the vaccine got covid-19. People who took the placebo got covid-19 a lot more. This is significant. 

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions